Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-05 WSSSPE1_report/WSSSPE1_summary_report.tex
quotes
Adding introduction to section 6.

%There were a number of papers categorized under the Communities banner and so
%two presenters each summarized half of the submissions in this area.
\todo{an intro to the whole section is needed}

@@

% What are the challenges?
Long-term sustainability of scientific software is often a significant
challenge and face radically different issues compared to utility software.
Many submissions reported that software is practically considered a `byproduct'
of the actual research. Others contended that the software was not the main
funded part of their research. A smaller codebase and fixed requirements result
in stability, ease of installation, and configurability.  Many such projects
mature and are treated as libraries to be used into larger systems such as some
of the utility software discussed in the previous section. While the software
can stay stable and require relatively low maintenance, the responsibility is
often on the shoulders of a very few developers who might not be specialists in
software development. Development tends to be linear and simplistic with a
limited scope to follow software best practices.

>>>>

%There were a number of papers categorized under the Communities banner and so
%two presenters each summarized half of the submissions in this area.
Across all talks and papers submitted, authors implicitly and explicitly
recognized the concept of ``communities'' as a driving and unifying force
within software projects.  Despite that, the actual nature of the communities,
the incentive structures and bind them together, the infrastructure they
utilize for communication and even the types of individuals that make up those
communities were different.  Some of these communities were primarily composed
of members of industry, some were funded and driven by individuals focused on
developing software as opposed to utilizing it, and others were primarily
composed of scientist practitioners, and were true communities of practice.

These varying structures and compositions result in differing modes of
interaction within communities, styles of development, and the structure of
planning for future development of software.  In this section, we summarize
the different types of communities in scientific software as well as the
resultant impact on sustainability and development of functionality.

@@

% What are the challenges?
Long-term sustainability of scientific software is often a significant
challenge and face radically different issues compared to utility software.
Many submissions reported that software is practically considered a ``byproduct"
of the actual research. Others contended that the software was not the main
funded part of their research. A smaller codebase and fixed requirements result
in stability, ease of installation, and configurability.  Many such projects
mature and are treated as libraries to be used into larger systems such as some
of the utility software discussed in the previous section. While the software
can stay stable and require relatively low maintenance, the responsibility is
often on the shoulders of a very few developers who might not be specialists in
software development. Development tends to be linear and simplistic with a
limited scope to follow software best practices.
